Problems with the characterisation of accreting Brown Dwarfs: A public grid of simulated SEDs and photometry

### 2010-04-19 BD_disc/old_paper.tex
New project
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2228 25464f21-6099-402c-b30a-640f1d40d38b

% If your system does not have the AMS fonts version 2.0 installed, then
% remove the useAMS option.  
% useAMS allows you to obtain upright Greek characters.
% e.g. \umu, \upi etc.  See the section on "Upright Greek characters" in
% this guide for further information.
%
% If you are using AMS 2.0 fonts, bold math letters/symbols are available
% at a larger range of sizes for NFSS release 1 and 2 (using \boldmath or
% preferably \bmath).
%
% The usenatbib command allows the use of Patrick Daly's natbib.sty for
% cross-referencing.
%
% If you wish to typeset the paper in Times font (if you do not have the
% PostScript Type 1 Computer Modern fonts you will need to do this to get
% smoother fonts in a PDF file) then uncomment the next line
% \usepackage{Times}
\usepackage{lscape}
\usepackage{graphicx}
\usepackage{subfigure}

@@

\title[**]{**}  \author[N.J.  Mayne and Tim
Harries]{N.J. Mayne$^{1}$\thanks{E-mail: nathan@astro.ex.ac.uk
    (NJM)} and Tim Harries$^{1}$\\
  $^{1}$ School of Physics,
  University of Exeter, Stocker Road, Exeter, EX4 4QL.\\
}
 \begin{document}

@@

\begin{abstract}
SORT OUT REFERENCE TO THE FILTER RESPONSES IN THE SEDS FIGURES say
discussed later
more about stellar photosphere overlapping with disc

FINESSE FIGURE SIZES

CHANGE ALL WORDING FOR NEW FIGURES** CHANGED AGE TO 1 MYR FOR SUB
FIGURES AND NEED TO ACKNOWLEDGE THROUGHT

CHOSE 10 MYRS AS PRE-MS CONTRACTION AGES ARE PROBABLY A  FACTOR OF 2 WRONG

ADD FILTERS TO OTHER SEDS FIGURES

Note about fitting objects in the next paper.

@@

@@

@@

The presence, or fraction (for populations) of associated discs is
itself also crucial in isolating the formation mechanisms of
BDs. Precise determination of disc parameters for BDs and comparison
with the TTS results may allow the differentiation of currently
competing theories of the BD formation mechanism. Essentially, a
mechanism similar to that seen in higher mass stars i.e. cloud
collapse and fragmentation will result in a similar fraction of discs
across the BD mass boundary, whereas a formation mechanism dominated
by dynamical interactions \citep[ejections, accretion disc stripping
from near collisions or tidal interactions etc.]{2003IAUS..211...27B}
will result in a markedly different disc parameters, such as fractions
of stars with discs or radial sizes of the discs themselves, across
the BD mass boundary. Fundamentally a different disc fraction for BD
stars to TTS would provide persuasive evidence of a different
formation mechanism across the two mass regimes, as would an increase
in the average radial size of discs, indicative of truncation for BD
objects. Therefore, to progress our understanding of the fundamental
issues of both the formation mechanism and emerging IMFs for BDs,
consideration of the effects of an associated circumstellar disc and
their subsequent ubiquity are essential.

@@

@@

@@

@@

@@

If disc-locking is present for BDD systems, as for TTS stars, one
would expect accretion of material from a magnetic truncation radius,
$R_{\rm T}$ \citep{2007prpl.conf..479B}.  \cite{2007prpl.conf..479B}
show that the magnetic truncation radius, $R_{\rm T}$, must be less
than the co-rotation radius, $R_{\rm co}$ (where the Keplerian angular
rotation rate of the disc is the same as the angular rotation rate of
the star), as material outside this radius will have too much angular
momentum to fall onto the star. This is implicitly adopted in the
proposed accretion mechanism, where the disc is magnetically
truncated, and therefore locked to the star. This magnetic truncation
would lead to an inner wall (and inner hole) of the accretion
disc. Evidence for an inner disc wall has been found from near-IR
excesses from the inner rims of Herbig Ae stars, with emission peaking
at 2-3$\mu$m \citep{2001ApJ...560..957D}. The actual shape and radial
position of this wall is affected by several factors, discussed
further in Sections \ref{dust_edge} and \ref{inner_edge}. The previous
work of \cite{1997AJ....114..288M} explored the effects of simple mass
accretion and the radius to the inner disc on the infrared colours of
CTTS, useing a prescribed range of inner disc
radii. \cite{1997AJ....114..288M} find that the scatter in infrared
colours for there parameter range, $R_{\rm inner}=$1$-$12$R_*$ and
$\dot{M_{\rm acc}}=$10$^{-9}$to 10$^{-5} M_{\odot}yr^{-1}$, matches
the scatter of the CTTS locus found in the data. The models of
\cite{1997AJ....114..288M} used physically flat discs however meaning
their results will not stand for discs with significant flaring. More
importantly for our work \cite{1997AJ....114..288M} find a correlation
of the infrared locus of models with the accretion and inner disc
location. Essentially, changes in the stellar flux levels and position
of the inner wall lead to temperature changes of this inner edge and
therefore changes in the balance of flux across the infrared
wavelength regime. As this study does not include the effects of
vertical hydrostatic equilibirum or even the study of flared discs one
would expect, perhaps, a more complex relationship between the inner
edge position and infrared excess in `real' systems. This means that
studies aiming to provide evidence of disc locking by comparing disc
indicators with rotation rates may include unforseen biases. As the
rotation rate will effect the co-rotation radius and therefore
truncate the disc at different radii, thereby enhancing or suppressing
any relationship of infrared excess (or disc presence) with rotation
rate.

In summary, several key areas in the theoretical studies of BDs hinge
critically on observational constraints, where the derived values are
heavily affected by disc and/or accretion presence.  Formation
mechanisms, IMFs, disc dissipation timescales, accretion rate--stellar
mass relationships and studies of disc-locking mechanisms using
correlations of rotation rates with infrared excesses, all require
derivations of either masses, ages or disc fractions (or a
combination), supplemented by other parameters. The consequences of
the disc structure for BD accretion discs, particularly vertical
structure, given the assumption of vertical hydrostatic equilibrium,
have already been discussed for a limited range of BD parameters in
\cite{2004MNRAS.351..607W}. Indeed \cite{2004MNRAS.351..607W} have
shown that BDD systems seen at higher inclinations can be confused
with CTTS systems, having similar colours and
magnitudes. Additionally, a benchmark model grid of spectra for star
and disc systems in the mass range 0.1$M_{\odot}<M_*<$50$M_{\odot}$
has previously been published in \cite{2006ApJS..167..256R}.

In this work we explore the effects of accretion and disc presence on
the derivation of the stellar parameters and explore the intrinsic
biases, or relationships, which will effect or inhibit conclusions on
mass to accretion rate or rotation to infrared excess correlations. We
have extended the similar work of \cite{2004MNRAS.351..607W} by
inclusion of a simple treatment of accretion and a sophisticated
implementation of dust sublimation (amongst other improvements). The
complete model grid, with derived photometry and isochrones, is
available online through our browsing tool
\footnote{http://www.astro.ex.ac.uk/research/bd\textunderscore disc}
which is described in Appendix \ref{website}. The aim of this online
database is to complement that of \cite{2006ApJS..167..256R}. Finally,
subsequent papers will introduce expansion to the model grid and an
online fitting tool.

The structure of this paper is as follows. In Section \ref{model} we
outline the model used, including an explanation of our physical
assumptions (Section \ref{physics}), a brief description of the
radiative transfer code (Section \ref{torus}), and an explanation of
the derivation of emergent quantities such as photometric magnitudes
(Section \ref{derived}). The ranges of the input variables are then
explained and, where possible, justified using observational
constraints in Section \ref{par_space}, with the total parameter space
covered shown in Table \ref{par_space_table}. Section
\ref{disc_struct} then contains analysis and description of the
resulting temperature and dust density structures of the discs. The
consequences of these changes in disc structure on the SEDs of these
objects is then discussed in Section \ref{sed_analysis}. Section
\ref{photometry} then details problems caused for the derivation of
the primary parameters age, mass and disc fractions from photometry
and the resulting relations built on these parameters; accretion rate
to stellar mass and rotational period to disc presence. Finally,
conclusions are drawn, with a summary, in Section
\ref{conclusions}.**REWRITE AT END**A CASE STUDY IS THEN INCLUDED AS
SECTION \ref{object}.**

@@

In this Section we detail the physical model adopted and assumptions
made (Section \ref{physics}), then explain key elements of the
radiative transfer code (Section \ref{torus}). Then we discuss the
derived values, such as broadband photometric magnitudes and colours
(in Section \ref{derived}). Some internal consistency checks are
detail in Appendix \ref{consistency}.

@@

@@

To model an accreting BDD system we must first model the underlying
photospheric flux. To provide a self-consistent set of model spectra
we have adopted a BD stellar interior and atmospheric model grid. We
have then constructed the total photospheric flux for any input value
of stellar age and mass by interpolating for surface gravity (log(g)),
effective temperature ($T_{\rm eff}$), radius ($R_*/R_{\odot}$) and
luminosity ($L_*/L_{\odot}$). These values were then used to
interpolate atmospheric spectra for flux (ergs s$^{-1}$cm$^{-2}
\AA^{-1}$) from 1200 to 2$\times$ 10$^7$ $\AA$ (or 0.12 to 2000
$\mu$m). These spectra were then smoothed using 200 logarithmically
spaced points. Careful inspection ensured that no spectral features
were removed during smoothing. The stellar interior models used for
this study are the `DUSTY00' models of \cite{2000ApJ...542..464C}
combined with the `AMES-Dusty', atmospheric models of
\cite{2000ApJ...542..464C}, which are all available online
\footnote{http://perso.ens-lyon.fr/france.allard/}. For our
approximate $T_{\rm eff}$ (effective temperature) range of $\approx
$3000K$<T_{\rm eff}<$1600K the AMES-Dusty atmospheres are most
applicable (2700K$>T_{\rm eff}>$1700K). We did try including dynamic
application of atmospheres based on the derived $T_{\rm eff}$,
i.e. using AMES-Cond for $T_{\rm eff}<$1700K, however this resulted in
large discontinuities between the model atmospheres and resulting
spectra. As this only affects stars at the very edge of our
temperature range, i.e. for the oldest and lowest mass objects (for
the AMES-Cond case), we have adopted the same model throughout.

@@

Secondly, we require a model of the accretion flux for the system. The
flux due to accretion was added onto the intrinsic photospheric flux
using the simple approximation of blackbody emission. The selected
accretion rate was used to derive an accretion luminosity ($L_{\rm acc}$),
where the material was modelled as falling from the disc inner edge
onto the surface of the star. $L_{\rm acc}$ is calculated according to
\ref{Lacc}, 
\begin{equation}
{L_{\rm acc}=\frac{GM_*\dot{M}}{R_*}\times (1-\frac{R_*}{R_{\rm inner}})},
\label{Lacc}
\end{equation}
where $M_*$ is the stellar mass, $\dot{M}$ the mass accretion rate,
$R_*$ the stellar radius and $R_{\rm inner}$ the radius of the disc inner
boundary.

The initial inner disc radius was set to be the co-rotation radius
(this is discussed in more detail in Section
\ref{disc_parameters}). During the radiative transfer simulations of
the disc the final inner dust disc radius may be extended past the
co-rotation radius. This is due to dust sublimation effects (see
Sections \ref{dust_edge} and \ref{inner_edge} for an
explanation). However, as the gas within the disc is not sublimated it
will always extend down to the co-rotation radius and we assume
accretion will still occur from this radius. A full discussion of the
differences between the initial and resulting final radii of the dust
disc can be found in Section \ref{inner_edge}. Once the accretion
luminosity was derived, an adopted areal coverage ($A$), over the
stellar surface, was used to calculate an effective temperature
($T_{\rm acc}$), for the accretion `hot' spot, where
\begin{equation}
{T_{\rm acc}=(\frac{L_{\rm acc}}{4\pi R^2_{*} \sigma \times A})^{\frac{1}{4}}}.
\label{Tacc}
\end{equation}

Finally, a blackbody flux distribution is generated at $T_{\rm acc}$
and added onto the intrinsic stellar photospheric flux. In general one
would expect this to be an overestimate of the accretion flux, as for
pre-MS BDs large convective zones are expected on the stellar surface,
and some of the accretion energy may act to further drive these
convective currents, meaning flux is lost. It is worth noting however
that observationally UV excesses are often used to recreate and then
subtract an assumed accretion flux using a blackbody flux curve,
essentially the reverse of this method.

@@

@@

For this work the disc outer edge was set at 300 AU, this was chosen
as a maximum size of the circumstellar
disc. \cite{2008A&A...486..877B} have shown that the disc outer radius
has little effect on the resulting SED. However, in our subsequent
paper we will include models for outer radii of 100 AU and plan to
extend this parameter range further to smaller values of the outer
radius in the future.

@@

The combined (accretion plus photosphere) SED is then used as a
boundary condition for the {\sc torus} radiative transfer code and
as a benchmark set of SEDs to model `naked' BD systems. The set of
`naked' photospheres (plus accretion) are diluted by the factor
$(R_*/Distance)^2$ to a distance of 10pc. The `negligibly' accreting,
`naked' stars can be used to produce absolute magnitude (and intrinsic
colour) isochrones for comparison. Whilst, the remainder are used to
model systems showing active accretion where no disc is detected. For
instance, \cite{2009arXiv0901.2603K} find 43 stars within their sample
are actively accreting whilst no disc is detected (out of a total
sample of 1253).

@@

\subsection{Radiative transfer code: {\sc torus}}
\label{torus}

@@

We have used the {\sc torus} radiative transfer code which is
described in \cite{2000MNRAS.315..722H}, including the subsequent
refinements i.e.  addition of Adaptive-Mesh-Refinement (AMR)
introduced in \cite{2004MNRAS.350..565H}. {\sc torus} uses the
criterion of \cite{1999A&A...344..282L} in a radiative equilibrium
algorithm. The simulation also calculates \textit{vertical hydrostatic
  equilibrium} and dust sublimation for the disc, similar to that
described in \cite{2007ApJ...661..374T}.

@@

We have also applied dust sublimation during our simulations. This
treatment is similar to that detailed in \cite{2007ApJ...661..374T},
but improved and generalised with respect to the applicable
geometries.

The practical implementation is as follows: An initial coarse
approximate temperature structure is found for the grid without dust,
only a fixed gas distribution. Then, first three full \textit{Lucy}
algorithm iterations are run for a grid with a maximum cell opacity of
0.1 (with dust added to this level). Three more iterations are then
performed with dust opacity limited to $\tau=$1 per cell. A further
three \textit{Lucy} algorithm iterations are then performed with the
full dust opacity, where dust is added until the gas-to-dust ratio is
achieved($\epsilon =$100). For each iteration (from the first to the
ninth) dust in cells which reach the sublimation temperature and
density is removed from the disc by setting the opacity to zero for
the cell. Once these nine iterations are complete, the grid is refined
around the disc photosphere, and the radiative transfer alogrithm is
iterated through to convergence. After this first convergence, a
vertical hydrostatic equilibrium algorithm is run, after which the
disc grid is unrefined and then re-refined around the disc
photosphere. This process continues for five hydrostatic equilibrium
calculations. Finally, once the hydrostatic equilibrium iteratins are
complete we perform a final dust sublimation run to remove any cells
where the dust temperature has moved above the sublimation temperature
and further refine the photosphere and inner edge of the disc (whilst
also unrefining cells where the opacity has reduced significantly and
high resolution is no longer required). In this model the dust
sublimation is effectively digital i.e. when the density and
temperature for sublimation is reached for a given cell the dust is
destroyed. We did attempt a more realistic model where the dust
sublimation causes the dust to be destroyed at an evanescently
decaying rate into the particular region. This was expected to lead to
a more realistic situation where the dust destruction propogates into
the disc material, and does not instantly destroy an entire
cell. However, this implementation lead to incremental sections of
dust being removed after each iteration, resulting in little change in
the overall density structure of the disc. This is perhaps interesting
for real systems, in that the dust sublimation should not be expected
to act as a digital `switch'.

>>>>

% If your system does not have the AMS fonts version 2.0 installed, then
% remove the useAMS option.  
% useAMS allows you to obtain upright Greek characters.
% e.g. \umu, \upi etc.  See the section on "Upright Greek characters" in
% this guide for further information.
%
% If you are using AMS 2.0 fonts, bold math letters/symbols are available
% at a larger range of sizes for NFSS release 1 and 2 (using \boldmath or
% preferably \bmath).
%
% The usenatbib command allows the use of Patrick Daly's natbib.sty for
% cross-referencing.
%
% If you wish to typeset the paper in Times font (if you do not have the
% PostScript Type 1 Computer Modern fonts you will need to do this to get
% smoother fonts in a PDF file) then uncomment the next line
% \usepackage{Times}
\usepackage{lscape}
\usepackage{graphicx}

@@

\title[Accreting Brown Dwarfs]{Problems with the characterisation of
  accreting Brown Dwarfs: A public grid of simulated SEDs and
  photometry}  \author[N.J.  Mayne and Tim
Harries]{N.J. Mayne$^{1}$\thanks{E-mail: nathan@astro.ex.ac.uk
    (NJM)} and Tim Harries$^{1}$\\
  $^{1}$ School of Physics,
  University of Exeter, Stocker Road, Exeter, EX4 4QL.\\
}
 \begin{document}

@@

\begin{abstract}

   1) All graphs of SEDS in log log units.

  1a) Add in the new data and write in using more masses.

  1b) Write in the extra inclinations-check from incs.lis.

  1c) Write in that using same points for photosphere.dat

  1) BOUY ET AL 2008, NEUGEBAUER 1984 AND SKRUTSKIE 2006 TRUNCATED AUTHOR LIST.

  2) CHANGE SYMBOLS OF NAKED STARS IN IR EXCESS AND IR EXCESS 2 TO
  MAKE CLEARER.

  3) REWRITE CONCLUSIONS WITH MORE REFERENCE TO
  \cite{2004MNRAS.351..607W}, \cite{2006A&A...452..245N} AND
  \cite{2006MNRAS.370L..10C}.

  4) REDO \& CHECK ALL FIGURES.

  5) COMPARE FINAL INNER RADIUS, AFTER SUBLIMATION, TO COROTATION
  RADIUS-HOT GAS STUFF.

  6) ADD IN DISCUSSION FOR MIPS AND CHECK ZEROPOINTS

  7) NEED A DISCUSSION OF THE DISC SCALEHEIGHT ETC (PROBABLY IN MODEL
  OR PAR SPACE): $\rho \propto R^{-\alpha}$, $\alpha =
  2.25$. $h\propto R^{\beta}$, $\beta = 1.25$. $\sigma = R^{(\beta
    -\alpha)}$, AND IS CONSERVED. $R_{\rm outer}=300$ AU. DISC SCALEHEIGHT
  AT 100 AU $H(100)=12.5$ AU

  8) ADD IN DISCUSSION OF THE TIME TAKEN FOR EACH MODEL AND
  SED-JUSTIFY SMALL NUMBER OF GRID. EACH MODEL (FOR ALL INCLINATIONS)
  =** HOURS OF SUPERCOMPUTER TIME ON ** CORES

  9) WHAT ABOUT EFFECT OF VERTICAL WALL FOR LONGER PERIOD OBJECTS (NO
  DUST SUBLIMATION) VERSUS SHORTER PERIOD AND CURVED WALL

  10) PUT IN INNER BOUNDARY PICTURES

  11) SORT OUT PLACEMENT OF CLEARPAGE.

  12) SORT RESULTS BIT ABOUT IRAC FLUX, THIS IS FROM REGIONS BEHIND
  THE INNER EDGE SO WHY IS THE TREND REVERSED, IT IS NOT COMING FROM
  THE OUTER FLARED REGIONS OF THE DISC BUT FROM JUST BEHIND THE PUFFED
  UP INNER EGDE. THEREFORE, FOR LONGER PERIOD OBJECTS THE INNER EDGE
  IS SMALLER AND THE DISC FLARED SO REGIONS JUST BEHIND CAN EMIT MORE
  FLUX. FOR SHORTER PERIOD OBJECTS, THE INNER EDGE IS PUFFED UP AND
  DOMINATES THE SED.

  13) TO PROVE THE POINT ABOVE PRODUCE AN IMAGE BATCH  FILE TO PRODUCE
  IMAGES FOR A TEST CASE I.E. CLOSE INNER EDGE DISTANT ONE ETC.

  14) Add in reference to the Mike Meyer paper about changes due to
  inner radius.

15) Reference to bouy et al 2008 showing that the outer radius has a
negligible effect

16) IMAGES.

17) Add in the Herschel filters.

18) Disc mass fraction.

19) IMF implications.

20) ADD IN RHO OPH 102 STUFF- and add in comparison to stars without accretion

21) Write in a bit about do we expect to see the correlation with
rotation rate.

22) Colours of 0.01Msol and 0.01Gyr models are very different.-REMOVE

23) Notes abnout isong 200 pts for SEDS naked in consisitency checks
used full resolution

24) NOTE ABOUT THE FACT THAT 0.01Msol STARS ARE OFF THE BOTTOM OF SOME
PLOTS AS THEY DIVE DOWN VERY ABRUPTLY.ASK ISABELLE ABOUT THIS

25) Add in comments to mention that due to numerics some models have
inner edges up to 100K above the 1700K mean dust sublimation temps.

26) Note about colour colour plots cut in magnitude for J and removed
the 0.01Msol stars from the naked

27) For detections add in piece about the overlapping stars i.e what
would be missed.

28) missing 0.01 in alpha graph also.

29) Adjust size of the smaller figures (i.e non-four) to fit.

30) Change argument around to say rotation period correlation not
present.

31) Plots of inner temp against dust sublimation inner radius etc

32) inclination against magnitude for different accretion rates.-one mass

33) Sot out JJK isochrone

34) Note about r=300 but outer radius doesnt matter largey see bouy
and vitell 2008 or something.

@@

\textbf{Motivation-why study discs around Brown dwarfs?}

@@

\textit{-Brown Dwarf formation mechanisms \& IMF.}

@@

The presence, or fraction (for populations) of associated discs is
itself also crucial in isolating the formation mechanisms of
BDs. Precise determination of disc parameters for BDs and comparison
with the TTS results may allow the differentiation of currently
competing theories of the BD formation mechanism. Essentially, a
mechanism similar to that seen in higher mass stars i.e. cloud
collapse and fragmentation will result in a similar fraction of discs
across the BD mass boundary, whereas a formation mechanism dominated
by dynamical interactions \citep[ejections, accretion disc stripping
from near collisions or tidal interactions etc.]{2003IAUS..211...27B}
will result in a markedly different disc parameters, such as fractions
of stars with discs or radial sizes of the discs themselves, across
the BD mass boundary. Fundamentally a different disc fraction for BD
stars to TTS would provide persuasive evidence of a different
formation mechanism across the two mass regimes, as would an increase
in the average radial size of discs, indicative of truncation for BD
objects.

Therefore, to progress our understanding of the fundamental issues of
both the formation mechanism and emerging IMFs for BDs, consideration
of the effects of an associated circumstellar disc and their
subsequent ubiquity are essential.

\textit{-Planet formation and disc dissipation timescales.}

@@

\textit{Mass dependent disc lifetimes.}

@@

\textit{-Accretion from an inner disc and clearing mechanisms.}

@@

\textit{-Accreting systems, relationship of stellar mass and accretion
  rate.}

@@

\textit{-Disc locking and star disc interaction: rotation.}

@@

If disc-locking is present for BDD systems, as for TTS stars, one
would expect accretion of material from a magnetic truncation radius,
$R_{\rm T}$ \citep{2007prpl.conf..479B}.  \cite{2007prpl.conf..479B}
show that the magnetic truncation radius, $R_{\rm T}$, must be less
than the co-rotation radius, $R_{\rm co}$ (where the Keplerian angular
rotation rate of the disc is the same as the angular rotation rate of
the star), as material outside this radius will have too much angular
momentum to fall onto the star. This is implicitly adopted in the
proposed accretion mechanism, where the disc is magnetically
truncated, and therefore locked to the star. This magnetic truncation
would lead to an inner wall (and inner hole) of the accretion
disc. Evidence for an inner disc wall has been found from near-IR
excesses from the inner rims of Herbig Ae stars, with emission peaking
at 2-3$\mu$m \citep{2001ApJ...560..957D}. The actual shape and radial
position of this wall is affected by several factors, discussed
further in Section \ref{physics}. The previous work of
\cite{1997AJ....114..288M} explored the effects of simple mass
accretion and the radius to the inner disc on the infrared colours of
CTTS, useing a prescribed range of inner disc
radii. \cite{1997AJ....114..288M} find that the scatter in infrared
colours for there parameter range, $R_{\rm inner}=$1$-$12$R_*$ and
$\dot{M_{\rm acc}}=$10$^{-9}$to 10$^{-5} M_{\odot}yr^{-1}$, matches
the scatter of the CTTS locus found in the data. The models of
\cite{1997AJ....114..288M} used physically flat discs however meaning
their results will not stand for discs with significant flaring. More
importantly for our work \cite{1997AJ....114..288M} find a correlation
of the infrared locus of models with the accretion and inner disc
location. Essentially, changes in the stellar flux levels and position
of the inner wall lead to temperature changes of this inner edge and
therefore changes in the balance of flux across the infrared
wavelength regime. As this study does not include the effectos of
vertical hydrostatic equilibirum or even the study of flared discs one
would expect, perhaps, a more complex relationship between the inner
edge position and infrared excess in `real' systems. This means that
studies aiming to provide evidence of disc locking by comparing disc
indicators with rotation rates may include unforseen biases. As the
rotation rate will effect the co-rotation radius and therefore
truncate the disc at different radii, thereby enhancing or suppressing
any relationship of infrared excess (or disc presence) with rotation
rate.

\textbf{Aims.}

\textit{Key areas: Require, Mass, age and disc fractions}

In summary, several key areas in the theoretical studies of BDs hinge
critically on observational constraints, where the derived values are
heavily affected by disc and/or accretion presence.  Formation
mechanisms, IMFs, disc dissipation timescales, accretion rate--stellar
mass relationships and studies of disc-locking mechanisms using
correlations of rotation rates with infrared excesses, all require
derivations of either masses, ages or disc fractions (or a
combination), supplemented by other parameters.

\textit{Previous complementary studies}

The consequences of the disc structure for BD accretion discs,
particularly vertical structure, given the assumption of vertical
hydrostatic equilibrium, have already been discussed for a limited
range of BD parameters in \cite{2004MNRAS.351..607W}. Indeed
\cite{2004MNRAS.351..607W} have shown that BDD systems seen at higher
inclinations can be confused with CTTS systems,
having similar colours and magnitudes. Additionally, a benchmark model
grid of spectra for star and disc systems in the mass range
0.1$M_{\odot}<M_*<$50$M_{\odot}$ has previously been published in
\cite{2006ApJS..167..256R}.

\textit{What we are actually doing in the paper}

In this work we explore the effects of accretion and disc presence on
the derivation of the stellar parameters and explore the intrinsic
biases, or relationships, which will effect or inhibit conclusions on
mass to accretion rate or rotation to infrared excess correlations. We
have extended the similar work of \cite{2004MNRAS.351..607W} by
inclusion of a simple treatment of accretion and a sophisticated
implementation of dust sublimation (amongst other improvements). The
complete model grid, with derived photometry and isochrones, is
available online through our browsing tool
\footnote{http://www.astro.ex.ac.uk/research/bd\textunderscore disc}
which is described in Appendix \ref{website}. The aim of this online
database is to complement that of \cite{2006ApJS..167..256R}. Finally,
a subsequent paper will introduce expansion to the model grid and an
online fitting tool.**

\textbf{Structure.}

**REDO THIS AT THE VERY END**

The structure of this paper is as follows. In Section \ref{model} we
outline the model used, including an explanation of our physical
assumptions (Section \ref{physics}), a brief description of the
radiative transfer code (Section \ref{torus}), and an explanation of
the derivation of emergent quantities such as photometric magnitudes
(Section \ref{derived}). The ranges of the input variables are then
explained and, where possible, justified using observational
constraints in Section \ref{par_space}. Section \ref{results} contains
the analysis and discussion of the main results of the model grid,
first discussing the effects on the derivations of the primary
parameters (Section \ref{derive_primary}), mass and age, and then the
effects on disc fraction derivation (Section \ref{derive_frac}).
Finally, conclusions are drawn, with a summary, in Section
\ref{conclusions}.

@@

In this Section we detail the physical model adopted and assumptions
made (Section \ref{physics}), then explain key elements of the
radiative transfer code (Section \ref{torus}). Then we discuss the
derived values, such as broadband photometric magnitudes and colours
(in Section \ref{derived}). Finally, we detail some internal
consistency checks (Section \ref{checks}).

@@

This section details some of the key physical assumptions made to
generate our model grid.

@@

\textit{-Model \& interpolation.}

In order to model an accreting BDD system we must first model the
underlying photospheric flux. To provide a self-consistent set of
model spectra we have adopted a BD stellar interior and atmospheric
model grid. We have then constructed the total photospheric flux for
any input value of stellar age and mass by interpolating for surface
gravity (log(g)), effective temperature ($T_{\rm eff}$), radius
($R_*/R_{\odot}$) and luminosity ($L_*/L_{\odot}$). These values were
then used to interpolate atmospheric spectra for flux (ergs
s$^{-1}$cm$^{-2} \AA^{-1}$) from 1200 to 2$\times$ 10$^7$ $\AA$ (or
0.12 to 2000 $\mu$m). These spectra were then smoothed using 200
logarithmically spaced points. Careful inspection ensured that no
spectral features were removed during smoothing. The stellar interior
models used for this study are the `DUSTY00' models of
\cite{2000ApJ...542..464C} combined with the `AMES-Dusty', atmospheric
models of \cite{2000ApJ...542..464C}, which are all available online
\footnote{http://perso.ens-lyon.fr/france.allard/}. For our
approximate $T_{\rm eff}$ (effective temperature) range of $\approx
$3000K$<T_{\rm eff}<$1600K the AMES-Dusty atmospheres are most
applicable (2700K$>T_{\rm eff}>$1700K). We did try including dynamic
application of atmospheres based on the derived $T_{\rm eff}$,
i.e. using AMES-Cond for $T_{\rm eff}<$1700K, however this resulted in
large discontinuities between the model atmospheres and resulting
spectra. As this only affects stars at the very edge of our
temperature range, i.e. for the oldest and lowest mass objects (for
the AMES-Cond case), we have adopted the same model throughout.

@@

\textit{-Blackbody flux given, Equation: Rinner, Coverage \& Rate.}

Secondly, we require a model of the accretion flux for the system. The
flux due to accretion was added onto the intrinsic photospheric flux
using the simple approximation of blackbody emission. The selected
accretion rate was used to derive an accretion luminosity ($L_{\rm acc}$),
where the material was modelled as falling from the disc inner edge
onto the surface of the star. $L_{\rm acc}$ is calculated acoording to
\ref{Lacc}, 
\begin{equation}
{L_{\rm acc}=\frac{GM_*\dot{M}}{R_*}\times (1-\frac{R_*}{R_{\rm inner}})},
\label{Lacc}
\end{equation}
where $M_*$ is the stellar mass, $\dot{M}$ the mass accretion rate,
$R_*$ the stellar radius and $R_{\rm inner}$ the radius of the disc inner
boundary.

The initial inner disc radius was set to be the co-rotation radius
(this is discussed in more detail in Section
\ref{disc_parameters}). During the radiative transfer simulations of
the disc the final inner dust disc radius may be extended past the
co-rotation radius. This is due to dust sublimation effects (see
Sections \ref{disc_parameters} and \ref{dust_edge} for
explanation). However, as the gas within the disc is not sublimated it
will always extend down to the co-rotation radius and we assume
accretion will still occur from this radius. A full discussion of the
differences between the initial and resulting final radii of the dust
disc can be found in Section \ref{inner_edge_position}. Once the
accretion luminosity was derived, an adopted areal coverage ($A$),
over the stellar surface, was used to calculate an effective
temperature ($T_{\rm acc}$), for the accretion `hot' spot, where
\begin{equation}
{T_{\rm acc}=\frac{L_{\rm acc}}{(4\pi R^2_{*} \sigma \times A)^{\frac{1}{4}}}}.
\label{Tacc}
\end{equation}

Finally, a blackbody flux distribution is generated at $T_{\rm acc}$
and added onto the intrinsic stellar photospheric flux. In general one
would expect this to be an overestimate of the accretion flux, as for
pre-MS BDs large convective zones are expected on the stellar surface,
and some of the accretion energy may act to further drive these
convective currents, meaning flux is lost. It is worth noting however
that observationally UV excesses are often used to recreate and then
subtract an assumed accretion flux using a blackbody flux curve.

@@

\textit{-Inner disc boundary, Co-rotation, i.e. magnetically
  truncated: rotational period.}

@@

\textit{-Disc outer edge.}

For this work the disc outer edge was set at 300 AU, this was chosen
as a maximum size of the circumstellar
disc. \cite{2008A&A...486..877B} have shown that the disc outer radius
has little effect on the resulting SED. However, in our subsequent
paper we have included models for outer radii of 100 AU and plan to
extend this parameter range further to smaller values of the outer
radius in the future.

@@

\textit{Naked and TORUS versions.}

The combined (accretion plus photosphere) SED is then used as a
boundary condition for the TORUS radiative transfer code and as a
benchmark set of SEDs to model `naked' BD systems. The set of `naked'
photospheres (plus accretion) are diluted by the factor
$(R_*/Distance)^2$ to a distance of 10pc. The `negligibly' accreting,
`naked' stars can be used to produce absolute magnitude (and intrinsic
colour) isochrones for comparison. Whilst, the remainder are used to
model systems showing active accretion where no disc is detected. For
instance, \cite{2009arXiv0901.2603K} find 43 stars within their sample
are actively accreting whilst no disc is detected (out of a total
sample of 1253).

@@

\subsection{Radiative transfer code: TORUS}
\label{torus}

@@

\textbf{TORUS.}

\textit{Radiative transfer code.}

We have used the TORUS radiative transfer code which is described in
\cite{2000MNRAS.315..722H}, including the subsequent refinements i.e.
addition of Adaptive-Mesh-Refinement (AMR) introduced in
\cite{2004MNRAS.350..565H}. TORUS uses the criterion of
\cite{1999A&A...344..282L} in a radiative equilibrium algorithm. The
simulation also calculates \textit{vertical hydrostatic equilibrium}
and dust sublimation for the disc, similar to that described in
\cite{2007ApJ...661..374T}.

@@

\textit{-Dust sublimation}

In addition, we have also applied dust sublimation during our
simulations. This treatment is similar to that
detailed in \cite{2007ApJ...661..374T}, but improved and generalised
with respect to the applicable geometries.

The practical implementation is as follows: An initial coarse
approximate temperature structure is found for the grid without dust,
only a fixed gas distribution. Then, first three full \textit{Lucy}
algorithm iterations are run for a grid with a maximum cell opacity of
0.1 (with dust added to this level). Three more iterations are then
performed with dust opacity limited to $\tau=$1 per cell. A further
three \textit{Lucy} algorithm iterations are then performed with the
full dust opacity, where dust is added until the gas-to-dust ratio is
achieved($\epsilon =$100). For each iteration (from the first to the
ninth) dust in cells which reach the sublimation temperature and
density is removed from the disc by setting the opacity to zero for
the cell. Once these nine iterations are complete, the grid is refined
around the disc photosphere, and the radiative transfer alogrithm is
iterated through to convergence. After this first convergence, a
vertical hydrostatic equilibrium algorithm is run, after which the
disc grid is unrefined and then re-refined around the disc
photosphere. This process continues for five hydrostatic equilibrium
calculations. In this model the dust sublimation is effectively
digital i.e. when the density and temperature for sublimation is
reached for a given cell the dust is destroyed. We did attempt a more
realistic model where the dust sublimation causes the dust to be
destroyed at an evanescently decaying rate into the particular
region. This was expected to lead to a more realistic situation where
the dust destruction propogates into the disc material, and does not
instantly destroy an entire region. However, this implementation lead
to incremental sections of dust being removed after each iteration,
resulting in little change in the overall density structure of the
disc. This is perhaps interesting for real systems, in that the dust
sublimation should not be expected to act as a digital `switch'.

\textit{Analytic disc structures}

We have also modeled the same grid without using vertical hydrostatic
equilibrium and, instead, adopting analytic solutions for the
structure of the disc. These models are included in our subsequent
paper describing the fitting routine and will be analysed in more
detail in a later work.

\textit{-Inner disc location}
